# Data Visualizations
## Building Energy Signature
### Task
Often the time series data from the outside temperature and the energy data are available in separate formats and time intervals. Therefore the following example imports and aggregates them separately.

### Solution
```{r energysignature ggplot, warning=FALSE, message=FALSE, collapse = TRUE}
library(ggplot2)
library(plotly)
library(dplyr)
library(redutils)
library(lubridate)

# load time series data and aggregate daily mean values
dfOutsideTemp <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralOutsideTemp.csv",
                          stringsAsFactors=FALSE, sep =";")
dfOutsideTemp$time <- parse_date_time(dfOutsideTemp$time, "YmdHMS", tz = "Europe/Zurich")
dfOutsideTemp$day <- as.Date(cut(dfOutsideTemp$time, breaks = "day"))
dfOutsideTemp <- dfOutsideTemp %>% group_by(day) %>% mutate(tempMean = mean(centralOutsideTemp)) %>% ungroup()
dfOutsideTemp <- dfOutsideTemp %>% select(day, tempMean) %>% unique() %>% na.omit()

dfHeatEnergy <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralHeating.csv",
                         stringsAsFactors=FALSE, sep =";")

dfHeatEnergy <- dfHeatEnergy %>% select(time, energyHeatingMeter) %>% na.omit()
dfHeatEnergy$time <- parse_date_time(dfHeatEnergy$time, "YmdHMS", tz = "Europe/Zurich")
dfHeatEnergy$day <- as.Date(cut(dfHeatEnergy$time, breaks = "day"))
dfHeatEnergy <- dfHeatEnergy %>% group_by(day) %>% mutate(energyMax = max(energyHeatingMeter)) %>% ungroup()
dfHeatEnergy <- dfHeatEnergy %>% select(day, energyMax) %>% unique() %>% na.omit()
dfHeatEnergy <- dfHeatEnergy %>% mutate(energyCons = energyMax - lag(energyMax)) %>% select(-energyMax) %>% na.omit()

# merge the data in a tidy format
df <- merge(dfOutsideTemp, dfHeatEnergy, by = "day")

# calculate season
df <- df %>% mutate(season = redutils::season(df$day))

# static chart with ggplot
p <- ggplot2::ggplot(df) +
  ggplot2::geom_point(aes(x = tempMean,
                          y = energyCons, color=season,
                          text = paste("</br>Date:  ", as.Date(df$day),
                                       "</br>Temp: ", round(df$tempMean, digits = 1), "\u00B0C",
                                       "</br>Energy: ", round(df$energyCons, digits = 0), "kWh/d",
                                       "</br>Season: ", df$season))
                      ) +
  ggtitle("Building Energy Signature") +
        theme_minimal() +
        theme(
          legend.position="none",
          plot.title = element_text(hjust = 0.5)
        )
p
```

Add the following part to your script to make the chart above interactive:

```{r energysignature plotly, warning=FALSE, message=FALSE, collapse = TRUE}
# continuation from upper ggplot code section
plotly::ggplotly(p, tooltip = c("text")) %>%
  layout(xaxis = list(title = "Outside temperature (\u00B0C)",
                      range = c(min(-5,min(df$tempMean)), max(35,max(df$tempMean))), zeroline = F),
         yaxis = list(title = "Daily energy consumption (kWh/d)",
                      range = c(-5, max(df$energyCons) + 10)),
         showlegend = TRUE
         ) %>%
  plotly::config(displayModeBar = FALSE, displaylogo = FALSE)
```

### Discussion
blabla

### See Also

## Mollier hx Diagram
### Task
Create a mollier h-x diagram with scatter plot of temperature- and humidity time series data including a comfort zone.

### Solution
```{r mollier hx, warning=FALSE, message=FALSE, collapse = TRUE}
library(redutils)
library(dplyr)
library(r2d3)
library(lubridate)

# read data
data <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatTempHum.csv",
                          stringsAsFactors=FALSE, sep =";")

# select temperature and humidity and remove empty cells
data <- data %>% select(time, FlatA_Temp, FlatA_Hum) %>% na.omit()

# create column with day for later grouping
data$time <- parse_date_time(data$time, "YmdHMS", tz = "Europe/Zurich")
data$day <- as.Date(cut(data$time, breaks = "day"))

# calculate daily mean of temperature and humidity
data <- data %>% group_by(day) %>% mutate(tempMean = mean(as.numeric(FlatA_Temp))) %>% ungroup()
data <- data %>% group_by(day) %>% mutate(humMean = mean(as.numeric(FlatA_Hum))) %>% ungroup()

# shrink down to daily values
data <- data %>% select(day, tempMean, humMean) %>% unique() %>% na.omit()

# plot mollier hx diagram
redutils::mollierHxDiagram(data)
```

### Discussion
blabla

### See Also
