# Data Wrangling
## Add Metadata for later filtering

Firstly we have to load a dataset into a dataframe:
```{r metadata dataset, warning=FALSE, message=FALSE, collapse = FALSE}
# load data set
df <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralOutsideTemp.csv",
               stringsAsFactors=FALSE,
               sep =";")

```

### Year, Month, Day, Day of Week
To group, filter and aggregate data we need to have a the date splitted up in day, month and year separately:
```{r metadata date, warning=FALSE, message=FALSE, collapse = FALSE}
library(dplyr)
library(lubridate)

df$time <- parse_date_time(df$time, "YmdHMS", tz = "Europe/Zurich")
df$year <- as.Date(cut(df$time, breaks = "year"))
df$month <- as.Date(cut(df$time, breaks = "month"))
df$day <- as.Date(cut(df$time, breaks = "day"))
df$weekday <- weekdays(df$time)
```

This code first parses the timestamp with a specific timezone. Then three columns are added.

Please note that the month also contains the year and a day. This is useful for a later step where you can group the series afterwards. 

\newpage

```{r metadata date output, warning=FALSE, message=FALSE, collapse = FALSE}
head(df,2)
tail(df,2)
```

### Season of Year
For some analyses it is useful to color single points of a scatterplot according to the season. For this we need to have the season in a separate column: 

```{r metadata season1, warning=FALSE, message=FALSE, collapse = FALSE}
# install redutils library
# devtools::install_github("retomarek/redutils", ref = "master")

# get season from a date
redutils::season(as.Date("2019-04-01"))
```

If you want to change the language, you can give the function dedicated names for the season:

```{r metadata season2, warning=FALSE, message=FALSE, collapse = FALSE}
redutils::season(as.Date("2019-04-01"),
                 c("Winter","Frühling","Sommer","Herbst"))
```

To apply this function to a whole dataframe we can use the dplyr mutate function. The code below creates a new column named "season":

```{r metadata season3, warning=FALSE, message=FALSE, collapse = FALSE}
# apply it for a data frame
df <- dplyr::mutate(df, season = redutils::season(df$time))
```

```{r metadata season3 output, warning=FALSE, message=FALSE, collapse = FALSE}
head(df,2)
tail(df,2)
```
## Data Frames
Firstly we have to load a dataset into a dataframe:
```{r metadata dataframe, warning=FALSE, message=FALSE, collapse = FALSE}
# load data set
df <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatTempHum.csv",
               stringsAsFactors=FALSE,
               sep =";")
```
### Change Row Names
```{r change row names, warning=FALSE, message=FALSE, collapse = FALSE}
# Print the header and the first line
head(df, 1)

# rename columns and print the header and the first line
names(df) <- c("timestamp","Hum_A","Temp_A","Hum_B","Temp_B","Hum_C","Temp_C","Hum_D", "Temp_D")
head(df, 1)
```
### Wide to Long
```{r rbasics wide2long, warning=FALSE, message=FALSE, collapse = FALSE}

# create a copy of the dataframe and print the header and the first five line
head(df, 5)

# convert wide to long format
df.long <- as.data.frame(tidyr::pivot_longer(df,
                                             cols = -timestamp,
                                             names_to = "sensor",
                                             values_to = "value",
                                             values_drop_na = TRUE)
                         )

# long format
head(df.long, 16)
```
### Long to Wide
```{r rbasics long2wide, warning=FALSE, message=FALSE, collapse = FALSE}
# long format
head(df.long)

# convert long table into wide table
df.wide <- as.data.frame(tidyr::pivot_wider(df.long,
                                            names_from = "sensor",
                                            values_from = "value")
                         )

# wide format
head(df.wide)
```
### Merge two Dataframes {#DATAWRANGLING-MERGE-TWO-DATAFRAMES}
```{r dataframe merge, warning=FALSE, message=FALSE, collapse = FALSE}
library(dplyr)
library(lubridate)

# read file one and parse dates
dfOutsideTemp <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralOutsideTemp.csv",
                          stringsAsFactors=FALSE,
                          sep =";")

dfOutsideTemp$time <- parse_date_time(dfOutsideTemp$time,
                                      orders = "YmdHMS",
                                      tz = "Europe/Zurich")

# read file two and parse dates
dfFlatTempHum <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatTempHum.csv",
                          stringsAsFactors=FALSE, sep =";")

dfFlatTempHum$time <- parse_date_time(dfFlatTempHum$time,
                                      order = "YmdHMS",
                                      tz = "Europe/Zurich")

# merge the two files into a new data frame and keep only rows where all values are available
df <- merge(dfOutsideTemp, dfFlatTempHum, by = "time") %>% na.omit()

```
