# Data Wrangling
## Add Metadata for later filtering

Firstly we have to load the required libraries and import a dataset into a dataframe:
```{r metadata dataset, warning=FALSE, message=FALSE, collapse = TRUE}

# load data set
data <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralOutsideTemp.csv",
               stringsAsFactors=FALSE, sep =";")

```

### Add Year, Month, Day
To group, filter and aggregate data we need to have a the date splitted up in day, month and year separately:
```{r metadata date, warning=FALSE, message=FALSE, collapse = TRUE}
library(dplyr)
library(lubridate)

df <- data

df$time <- parse_date_time(df$time, "YmdHMS", tz = "Europe/Zurich")
df$year <- as.Date(cut(df$time, breaks = "year"))
df$month <- as.Date(cut(df$time, breaks = "month"))
df$day <- as.Date(cut(df$time, breaks = "day"))
```

This code first parses the timestamp with a specific timezone. Then three columns are added.

Please note that the month also contains the year and a day. This is useful for a later step where you can group the series afterwards. 

```{r metadata date output, warning=FALSE, message=FALSE, collapse = TRUE}
head(df,2)
tail(df,2)
```

### Add Season of Year
For some analyses it is useful to color single points of a scatterplot according to the season. For this we need to have the season in a separate column: 

```{r metadata season1}
# install redutils library
# devtools::install_github("retomarek/redutils", ref = "master")

# get season from a date
redutils::season(as.Date("2019-04-01"))
```

If you want to change the language, you can give the function dedicated names for the season:

```{r metadata season2}
redutils::season(as.Date("2019-04-01"), c("Winter","Frühling","Sommer","Herbst"))
```

To apply this function to a whole dataframe we can use the dplyr mutate function:

```{r metadata season3}

df <- data

# apply it for a data frame
df <- dplyr::mutate(df, season = redutils::season(df$time))
```

```{r metadata season3 output, warning=FALSE, message=FALSE, collapse = TRUE}
head(df,2)
tail(df,2)
```
