# Data Visualizations
## Building Energy Signature
### Basis

* Two separate csv files with time series data from the outside temperature and the energy data with unaligned time intervals 

* Energy consumption time series from a energy meter with steadily increasing meter values

### Task
You want to create a scatter plot with

* the daily mean outside temperature on the x-axis

* the daily energy consumption on the y-axis

* points colored according to season


### Solution
After reading in the two time series the data has to get aggregated per day and then merged. Note that during the aggregation of the energy data you have to calculate the daily conspumption from the steadiliy increasing meter values as well.

Create a new script, copy/paste the following code and run it:

```{r energysignature ggplot, warning=FALSE, message=FALSE, collapse = FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
library(redutils)
library(lubridate)

# load time series data and aggregate daily mean values
dfOutsideTemp <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralOutsideTemp.csv",
                          stringsAsFactors=FALSE, sep =";")
dfOutsideTemp$time <- parse_date_time(dfOutsideTemp$time, "YmdHMS", tz = "Europe/Zurich")
dfOutsideTemp$day <- as.Date(cut(dfOutsideTemp$time, breaks = "day"))
dfOutsideTemp <- dfOutsideTemp %>% group_by(day) %>% mutate(tempMean = mean(centralOutsideTemp)) %>% ungroup()
dfOutsideTemp <- dfOutsideTemp %>% select(day, tempMean) %>% unique() %>% na.omit()

dfHeatEnergy <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralHeating.csv",
                         stringsAsFactors=FALSE, sep =";")

dfHeatEnergy <- dfHeatEnergy %>% select(time, energyHeatingMeter) %>% na.omit()
dfHeatEnergy$time <- parse_date_time(dfHeatEnergy$time, "YmdHMS", tz = "Europe/Zurich")
dfHeatEnergy$day <- as.Date(cut(dfHeatEnergy$time, breaks = "day"))
dfHeatEnergy <- dfHeatEnergy %>% group_by(day) %>% mutate(energyMax = max(energyHeatingMeter)) %>% ungroup()
dfHeatEnergy <- dfHeatEnergy %>% select(day, energyMax) %>% unique() %>% na.omit()
dfHeatEnergy <- dfHeatEnergy %>% mutate(energyCons = energyMax - lag(energyMax)) %>% select(-energyMax) %>% na.omit()

# merge the data in a tidy format
df <- merge(dfOutsideTemp, dfHeatEnergy, by = "day")

# calculate season
df <- df %>% mutate(season = redutils::season(df$day))

# static chart with ggplot
p <- ggplot2::ggplot(df) +
  ggplot2::geom_point(aes(x = tempMean,
                          y = energyCons, color=season,
                          text = paste("</br>Date:  ", as.Date(df$day),
                                       "</br>Temp: ", round(df$tempMean, digits = 1), "\u00B0C",
                                       "</br>Energy: ", round(df$energyCons, digits = 0), "kWh/d",
                                       "</br>Season: ", df$season))
                      ) +
  ggtitle("Building Energy Signature") +
        theme_minimal() +
        theme(
          legend.position="none",
          plot.title = element_text(hjust = 0.5)
        )
p
```

Add the following part to your script to make the chart above interactive:

```{r energysignature plotly, warning=FALSE, message=FALSE, collapse = FALSE}
# continuation from upper ggplot code section
plotly::ggplotly(p, tooltip = c("text")) %>%
  layout(xaxis = list(title = "Outside temperature (\u00B0C)",
                      range = c(min(-5,min(df$tempMean)), max(35,max(df$tempMean))), zeroline = F),
         yaxis = list(title = "Daily energy consumption (kWh/d)",
                      range = c(-5, max(df$energyCons) + 10)),
         showlegend = TRUE
         ) %>%
  plotly::config(displayModeBar = FALSE, displaylogo = FALSE)
```

## Mollier hx Diagram
### Basis

* A csv file with time series from multiple temperature and humidity sensors in °C and %rH

### Task
You want to plot a mollier h-x diagram with

* scatter plot of temperature- and humidity sensor data (mean values per day)

* points colored according to season

* comfort zone

### Solution
The sensor data is not in a constant intervall and not yet aggregated. So after reading in the time series the data has to get filtered and aggregated per day.

Finally use the plot function `mollierHxDiagram` from the `redutils` package (R Energy Data Utilities) which you can install as followed:

```{r eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("hslu-ige-laes/redutils")
```
Create a new script, copy/paste the following code and run it:

```{r mollier hx, warning=FALSE, message=FALSE, collapse = FALSE}
library(redutils)
library(dplyr)
library(r2d3)
library(lubridate)

# read data
data <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatTempHum.csv",
                          stringsAsFactors=FALSE, sep =";")

# select temperature and humidity and remove empty cells
data <- data %>% select(time, FlatA_Temp, FlatA_Hum) %>% na.omit()

# create column with day for later grouping
data$time <- parse_date_time(data$time, "YmdHMS", tz = "Europe/Zurich")
data$day <- as.Date(cut(data$time, breaks = "day"))

# calculate daily mean of temperature and humidity
data <- data %>% group_by(day) %>% mutate(tempMean = mean(as.numeric(FlatA_Temp))) %>% ungroup()
data <- data %>% group_by(day) %>% mutate(humMean = mean(as.numeric(FlatA_Hum))) %>% ungroup()

# shrink down to daily values and remove rows with empty values
data <- data %>% select(day, tempMean, humMean) %>% unique() %>% na.omit()

# plot mollier hx diagram
redutils::mollierHxDiagram(data)
```

### Discussion
The diagram is based on D3 and packaged into the package `redutils`. The original D3 source with a html integration you can find here: <a href="https://github.com/hslu-ige-laes/d3-mollierhx" target="_blank">https://github.com/hslu-ige-laes/d3-mollierhx</a>

### See Also
If your two time series are in separate files, you must first read them in separately and then merge them into one data frame. See chapter \@ref(DATAWRANGLING-MERGE-TWO-DATAFRAMES)
