## Mollier hx Diagram
### Task
You want to plot a mollier h-x diagram with

* scatter plot of temperature- and humidity sensor data (mean values per day)

* points colored according to season

* comfort zone

### Basis

* A csv file with time series from multiple temperature and humidity sensors in °C and %rH

### Solution
The sensor data is not in a constant intervall and not yet aggregated. So after reading in the time series the data has to get filtered and aggregated per day.

Finally use the plot function `mollierHxDiagram` from the `redutils` package (R Energy Data Utilities).
If you have not yet installed this package, proceed as follows:

```{r eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("hslu-ige-laes/redutils")
```
Create a new script, copy/paste the following code and run it:

```{r mollier hx, warning=FALSE, message=FALSE, collapse = FALSE}
library(redutils)
library(dplyr)
library(lubridate)

# read and print data
data <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatTempHum.csv",
                 stringsAsFactors=FALSE,
                 sep =";")

# select temperature and humidity and remove empty cells
data <- data %>% select(time, FlatA_Temp, FlatA_Hum) %>% na.omit()

# create column with day for later grouping
data$time <- parse_date_time(data$time, "YmdHMS", tz = "Europe/Zurich")
data$day <- as.Date(cut(data$time, breaks = "day"))

# calculate daily mean of temperature and humidity
data <- data %>%
  group_by(day) %>%
  mutate(tempMean = mean(as.numeric(FlatA_Temp)),
         humMean = mean(as.numeric(FlatA_Hum))
         ) %>%
  ungroup()

# shrink down to daily values and remove rows with empty values
data <- data %>% select(day, tempMean, humMean) %>% unique() %>% na.omit()

# plot mollier hx diagram
plotMollierHx(data)
```

