# Getting started
Before we can start with the analysis described in the later chapters, we have to install "R".
Therefore the first two parts of this chapter get you up and running with downloading and installing the relevant software and packages.

This may seem laborious, but it is necessary and easier than it appears at first glance.

If you already have "R" and "R Studio" installed, please take a look at the [https://github.com/hslu-ige-laes/redutils](https://github.com/hslu-ige-laes/redutils){target="_blank"} package on github, which is frequently used throughout this book. Chapter \@ref(installationPackages) shows you how to install it.

Finally "R-beginners" find in \@ref(createFirstScript) an easy code example which creates a simple time series plot.

## Install R and R Studio {#installation}
Before we can start the first analysis, we have to install "R" and "RStudio". 

- "R" is a programming language used for statistical computing while "RStudio" provides a graphical user interface
- "R" may be used without "RStudio", but "RStudio" may not be used without "R"
- Both, "R" and "RStudio" are free of charge and there are no licencse fees
- When you later make an analysis and visualizations, you only work in the graphical user interface "RStudio"

\newpage
### Download and Install R
<br>
<font size="5">Windows</font><br>

1. Open [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/){target="_blank"} and press the link "Download R..."
1. Run the downloaded installer file and follow the installation wizard

The wizard will install "R" into your `Program Files` folders and adds a shortcut in your Start menu. Note that you will need to have all necessary administration rights to install new software on your machine.

<br>
<font size="5">Mac OSX</font><br>
1. Open [https://cran.r-project.org/bin/macosx/](https://cran.r-project.org/bin/macosx/){target="_blank"} and download the latest `*.pkg` file
1. Run the downloaded installer file and follow the installation wizard

The installer allows you to customize your installation. However the default values will be suitable for most users.

<br>
<font size="5">Liux</font><br>
"R" is part of many Linux distributions, therefore you should check with your Linux package management system if it's already installed.

The CRAN website provides files to build "R" from source on Debian, Redhat, SUSE, and Ubuntu systems under the link “Download R for Linux”

- Open [https://cran.r-project.org/bin/linux/](https://cran.r-project.org/bin/linux/){target="_blank"} and then follow the directory trail to the version of Linux you wish to install R on top of

The exact installation procedure will vary depending on your Linux operating system. CRAN supports the process by grouping each set of source files with documentation or README files that explain how to install on your system.

### Download and Install RStudio
1. Open [https://rstudio.com/products/rstudio/download/](https://rstudio.com/products/rstudio/download/){target="_blank"} and download "RStudio Desktop Open Source"
1. Follow the on-screen instructions
1. Once you have installed "R Studio", you can run it like any other application by clicking the program icon 

\newpage
## Install required packages {#installationPackages}
Appendix \@ref(packages) gives you an introduction to what a package is and how to install it. Below are the packages used in this book and it is recommended to install them.

- Open "RStudio" just as you would any program, by clicking on its icon

- Copy the following code and paste it into your console (on the bottom left, right of the symbol `>`):
```{r install packages, eval=FALSE}
install.packages("devtools", "tidyverse", "plotly", "lubridate", "r2d3")
install_github("hslu-ige-laes/redutils")
```
- Press `Enter` or `Return`

The installation of the packages is now in progress and this may take a while, please be patient. In the meantime you can read in appendix \@ref(packages) what packages are in general and how they can be installed and later loaded into scripts.

\newpage
## Create your first R Script {#createFirstScript}
Finally, you have installed "R" and "RStudio" with some packages on your computer. Good, hopefully everything worked fine up to now.

Let's create the first script with a visualization:

- Open "RStudio" just as you would any program, by clicking on its icon

- Go to the menu on the top left and click to `File / New File / R Script`

- Copy the following code and paste it into your script:

```{r co2 intro code, eval=FALSE}
library(graphics)
plot(co2, ylab = "CO2 (ppm)", las = 1)
title(main = "Mauna Loa Atmospheric CO2 Concentration")
```

- select all by pressing `Ctrl + A`

- Thren run the code by pressing the `Run` Button or `Ctrl + Enter`


You should now get your first visualization:

```{r co2 intro graph, echo=FALSE}
library(graphics)
plot(co2, ylab = "CO2 (ppm)", las = 1)
title(main = "Mauna Loa Atmospheric CO2 Concentration")
```

> As you probably noticed, we did not load any data. The basic installation of "R" and some packages come with test data. So that is an easy way to test something. [The R Dataset Package](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html){target="_blank"} provides some preinstalled datasets, including the used "Mauna Loa Atmospheric CO2 Concentration" dataset. 
