# Getting started
## Install R and R Studio {#installation}
Before we can start the first analysis, we have to install "R" and "RStudio". This may seem laborious, but it is necessary and easier than it appears at first glance.

- "R" is a programming language used for statistical computing while "RStudio" provides a graphical user interface.
- "R" may be used without "RStudio", but "RStudio" may not be used without "R".
- Both, "R" and "RStudio" are free of charge and there are no licencse fees.
- When you later make an analysis and visualizations, you only work in the graphical user interface "RStudio".

### Download and Install R
#### Windows

1. Open [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/) and press the link "Download R..."
1. Run the downloaded installer file and follow the installation wizard

The wizard will install R into your "Program Files" folders and add a shortcut in your Start menu. Note that you will need to have all necessary administration rights to install new software on your machine.

#### Mac OSX
1. Open [https://cran.r-project.org/bin/macosx/](https://cran.r-project.org/bin/macosx/) and download the latest *.pkg file
1. Run the downloaded installer file and follow the installation wizard

The installer allows you to customize your installation. However the default values will be suitable for most users.

#### Linux
R is part of many Linux distributions, therefore you should check with your Linux package management system if it's already installed.

The CRAN website provides files to build R from source on Debian, Redhat, SUSE, and Ubuntu systems under the link “Download R for Linux”

- Open [https://cran.r-project.org/bin/linux/](https://cran.r-project.org/bin/linux/) and then follow the directory trail to the version of Linux you wish to install R on top of

The exact installation procedure will vary depending on your Linux operating system. CRAN supports the process by grouping each set of source files with documentation or README files that explain how to install on your system.

### Download and Install RStudio
R Studio is a development environment for R.

1. Open [https://rstudio.com/products/rstudio/download/](https://rstudio.com/products/rstudio/download/) and download "RStudio Desktop Open Source"
1. Follow the on-screen instructions
1. Once you have installed R Studio, you can run it like any other application via 

## Install required packages
Chapter \@ref(packages) gives you an introduction to what a package is and how to install it. Below are the packages we use in this book, and it is recommended to install them now.

- Open "RStudio" just as you would any program, by clicking on its icon.
- Copy the following code and paste it into your console right of the symbol `>`:
```{r install packages, eval=FALSE}
install.packages("devtools", "tidyverse", "plotly", "lubridate", "r2d3")
install_github("hslu-ige-laes/redutils")
```
- Press enter

The installation of the packages is now in progress and this may take a while, please be patient. In the meantime you can read in chapter \@ref(Packages) what packages are in detail and how they can be installed and then loaded into scripts.

## Create your first R Script and visualization
Finally, you have both "R" and "RStudio" with the first set of packages on your computer. Let's create the first visualization.

- Open "RStudio" just as you would any program, by clicking on its icon.
- Go to the menu on the top left and click to `File / New File / R Script`
- Copy the following code and paste it into your script:
```{r co2 intro code, eval=FALSE}
library(graphics)
plot(co2, ylab = "CO2 (ppm)", las = 1)
title(main = "Mauna Loa Atmospheric CO2 Concentration")
```
- select all by pressing `Ctrl + A`
- Thren run the code by pressing the `Run` Button or `Ctrl + Enter`
- You should now get your first visualization:
```{r co2 intro graph, echo=FALSE}
library(graphics)
plot(co2, ylab = "CO2 (ppm)", las = 1)
title(main = "Mauna Loa Atmospheric CO2 Concentration")
```

