## Examples
### Import csv file from GWF Relay W60 M-Bus Logger
Some data loggers have a cryptic data format which requires some data wrangling ahead before we can use the time series efficiently. Following an example of a M-Bus data loggger.

```{r w60 read csv, warning=FALSE, message=FALSE, collapse = FALSE}
library(tidyr)

# load csv file
df <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatHeatAndHotWater_GWF_RelayW60Logger.csv",
               stringsAsFactors=FALSE,
               sep =",")
head(df,23)
```

Please note
 * The Date and Time are in separate columns
 * The HCA-Unit value has a factor of 10, so the above HCA-value of 87 are in 10 liters Units and result in 87 * 10 = 870 liter
 * In the example "Adr" is the identification of the flat
 
We are only interested in Nr. 1 heating energy of the flat and Nr. 12 the impulse count of the additional meter which is the hot water of the flat.

Following the required steps to parse the data and bring them in a format we can later work with:

```{r w60 parsing, warning=FALSE, message=FALSE, collapse = FALSE}

# create timestamp out of column "Date" and "Time"
df <- df %>% dplyr::mutate(timestamp = paste0(df$Datum, " ", df$Zeit))
df$timestamp <- parse_date_time(df$timestamp,
                                order = "d.m.y H:M",
                                tz = "Europe/Zurich")


# select columns and rearrange
df <- df %>% select(timestamp, Adr, Nr., Wert, Einheit)

# filter out Adr. of interest
df <- df %>% filter(Nr. %in% c(1,12))

# rename columns
df <- df %>% dplyr::mutate(Nr. = ifelse(Nr. == 1, paste0("Adr",sprintf("%02.0f", Adr), "_energyHeat"), ifelse(Nr. == 12, paste0("Adr",sprintf("%02.0f", Adr), "_hotWater"), Nr.)))

# convert value to numeric
df$Wert <- as.numeric(df$Wert)

# multiply HCA-values by factor 10 to get liters and divide by 1000 to get m3
df <- df %>% dplyr::mutate(Wert = ifelse((Einheit == "HCA"), Wert * 10/1000, Wert))

df <- df %>% select(-Einheit, -Adr)

# convert long table into wide table
df.wide <- as.data.frame(pivot_wider(df,
                                     names_from = "Nr.",
                                     values_from = Wert,
                                     names_sep = "_")
)
```

This is the result:
```{r w60 result, warning=FALSE, message=FALSE, collapse = FALSE}
head(df.wide)
```

Finally saving the csv file:
```{r w60 save csv, eval=FALSE}
write.csv2(df.wide,
           file = "flatHeatAndHotWater.csv",
           row.names = FALSE)
```
