# Explorative Data Analysis
Explorative Data Analysis (EDA) is a technique based on the human characteristic of visual pattern recognition.

The purpose of EDA is simple: learn more about data by visualizing it in different ways.

> “Exploratory data analysis is graphical detective work.” - John W. Tukey, considered the founder of EDA

## Get Overview of Data
A first step is getting an overview of the whole data set and specific series of it.

### Load data
Load test data set in a data frame (e.g. from a csv-file):

```{r eda readcsv, echo=FALSE, results='hide'}
df <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralHeating.csv",
                         stringsAsFactors=FALSE,
                         sep =";")

df$time <- parse_date_time(df$time,
                           orders = "YmdHMS",
                           tz = "Europe/Zurich")
```

<br>

### Names
Show the column headers of a data frame:

```{r eda names}
names(df)
```

<br>

### Structure
Show the structure of the data frame:

```{r eda str}
str(df)
```

<br>

### Head/Tail
Show the first and last values:

```{r eda head}
head(df)
```

```{r eda tail, warning=FALSE, message=FALSE, collapse = FALSE}
tail(df)
```

> Note: if you want to show only the first three entries, you can type `head(df,3)`

<br>

### Five number summary

Reveals details about the distribution of the data:

```{r eda summary, warning=FALSE, message=FALSE, collapse = FALSE}
summary(df$supplyTempHeating)
```

<br>

## Quick Data Visualizations
Please refer to the [R Graphics Cookbook from Winston Chang - Chapter 2 "Quickly Exploring Data"](https://r-graphics.org/chapter-quick){target="_blank"}.

This is a great resource where you can find everything you need to know about creating basic plots.
Some plots of the part "Data Visualizations" are as well easy to create and can get used in the context of EDA.



```{r eda qda tsd, warning=FALSE, message=FALSE, collapse = TRUE}
library(dplyr)
library(lubridate)
# load csv file
df <- read.csv2("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/flatHeatAndHotWater.csv",
                stringsAsFactors=FALSE)
# filter flat
df <- df %>% select(timestamp, Adr01_energyHeat, Adr02_energyHeat, Adr03_energyHeat)
#df <- df %>% filter(timestamp > "2014-12-01")

df <- df %>% dplyr::mutate(FlatA = lead(Adr01_energyHeat) - Adr01_energyHeat)
df <- df %>% dplyr::mutate(FlatB = lead(Adr02_energyHeat) - Adr02_energyHeat)
df <- df %>% dplyr::mutate(FlatC = lead(Adr03_energyHeat) - Adr03_energyHeat)

# remove counter value column
df <- df %>% select(-Adr01_energyHeat, -Adr02_energyHeat, -Adr03_energyHeat) %>% na.omit()
df$timestamp <- parse_date_time(df$timestamp,
                                orders = "YmdHMS",
                                tz = "Europe/Zurich")

tsd <- ts(df %>% select(-timestamp), frequency = 12, start = min(year(df$timestamp)))
```

### Multiple time series plot
```{r eda qda simple ts, warning=FALSE, message=FALSE, collapse = FALSE}
# plot FlatA_Ele
plot(tsd)
```

### Multiple time series plot in one plot
```{r eda qda all ts, warning=FALSE, message=FALSE, collapse = FALSE}
# Plotting all in one frame
colours<-c("green3","red3","blue3")

plot(tsd,plot.type="single",ylab="(kWh)",col=colours)

# Legend
ltxt<-c("Flat A", "Flat B", "Flat C")
legend("topleft",lty=1,col=colours, legend=ltxt)
```

### Indexed time series plot
```{r eda qda indexed ts, warning=FALSE, message=FALSE, collapse = FALSE}
tsdPlot <- tsd

# Index timeseries
tsdPlot[,1]<-tsdPlot[,1]/tsdPlot[1,1]*100
tsdPlot[,2]<-tsdPlot[,2]/tsdPlot[1,2]*100
tsdPlot[,3]<-tsdPlot[,3]/tsdPlot[1,3]*100

# Plotting all in one frame
colours<-c("green3","red3","blue3")

plot(tsdPlot,plot.type="single",ylab="Index",col=colours)

# Legend
ltxt<-c("Flat A", "Flat B", "Flat C")
legend("topleft",lty=1,col=colours, legend=ltxt)

```
