# Explorative Data Analysis
Explorative Data Analysis (EDA) is a technique based on the human characteristic of visual pattern recognition. The purpose of EDA is simple: to learn more about data by visualizing it in different ways. John W. Tukey, considered the founder of exploratory data analysis, once wrote: 

> “Exploratory data analysis is graphical detective work.” - John W. Tukey

## Get Overview of Data
A first step is getting an overview of the whole data set and specific series of it.

### Load data
Load test data set in a data frame (e.g. from a csv-file)

```{r eda readcsv, echo=FALSE, results='hide'}
df <- read.csv("https://github.com/hslu-ige-laes/edar/raw/master/sampleData/centralHeating.csv",
                         stringsAsFactors=FALSE,
                         sep =";")

df$time <- parse_date_time(df$time,
                           orders = "YmdHMS",
                           tz = "Europe/Zurich")
```

<br>

### Names
Show the column headers of a data frame

```{r eda names}
names(df)
```

<br>

### Structure
Show the structure of the data frame

```{r eda str}
str(df)
```

<br>

### Head/Tail
Show the first and last values

```{r eda head}
head(df)
```

```{r eda tail, warning=FALSE, message=FALSE, collapse = FALSE}
tail(df)
```

> Note: if you want to show only the first three entries, you can type `head(df,3)`

<br>

### Five number summary

Reveals details like object classes

```{r eda summary, warning=FALSE, message=FALSE, collapse = FALSE}
summary(df$supplyTempHeating)
```

<br>

## Quickly Exploring Data
Please refer to the <a href="https://r-graphics.org/chapter-quick" target="_blank">R Graphics Cookbook from Winston Chang - Chapter 2 "Quickly Exploring Data"</a>.

This is a great resource where you can find everything you need to know about creating basic plots.
